---
title: "Project2"
output: html_document
date: "2023-11-14"
---

```{r}
library(dplyr)
library(ggplot2)
library(nycflights13)
```
```{r}
UA_flights <- flights %>% 
  filter(carrier=="UA")
UA_flights <- UA_flights %>%
  filter(!is.na(dep_delay))
UA_flights <- UA_flights %>%
  filter(!is.na(arr_delay))

```
```{r}
UA_flights<- UA_flights %>%
  mutate(netGain= dep_delay - arr_delay)
```
```{r}
summary(UA_flights$netGain)
hist(UA_flights$netGain)
```
```{r}
UA_flights<- UA_flights %>%
  mutate(late = ifelse(dep_delay > 0, 1, 0),
           very_late = ifelse(dep_delay > 30, 1, 0),
         not_late=ifelse(dep_delay<=0,1,0),
         not_verylate=ifelse(dep_delay<=30,1,0))
```
```{r}
ggplot(UA_flights, aes(x = netGain)) +
  geom_histogram(binwidth = 2) +
  facet_wrap(~ late) +
  labs(title = "NetGain Distribution for late flights")
ggplot(UA_flights, aes(x = netGain)) +
  geom_histogram(binwidth = 2) +
  facet_wrap(~ very_late) +
  labs(title = "NetGain Distribution for very late flights")
ggplot(UA_flights, aes(x = netGain)) +
  geom_histogram(binwidth = 2) +
  facet_wrap(~ not_late) +
  labs(title = "NetGain Distribution for flights with no delay")
ggplot(UA_flights, aes(x = netGain)) +
  geom_histogram(binwidth = 2) +
  facet_wrap(~ not_verylate) +
  labs(title = "NetGain Distribution for flights which are not very late")
```
```{r}
late1 <- UA_flights$late == 1
veryLate1 <- UA_flights$very_late==1
notLate1 <- UA_flights$not_late==1
notveryLate1 <- UA_flights$not_verylate==1
avGainLate<- mean(UA_flights$netGain[late1])
cat('Average gain for flights that are late: ',avGainLate,'\n')
avGainVeryLate<- mean(UA_flights$netGain[veryLate1])
cat('Average gain for flights that are very late: ',avGainVeryLate,'\n')
avGainNotLate<- mean(UA_flights$netGain[notLate1])
cat('Average gain for flights that are not late: ',avGainNotLate,'\n')
avGainNotVeryLate<- mean(UA_flights$netGain[notveryLate1])
cat('Average gain for flights that are not very late: ',avGainNotVeryLate,'\n')

```

```{r}
glimpse(UA_flights)
# Permutation Tests for late vs not late
observed <- avGainLate - avGainNotLate
N <- 10^4 - 1

result <- numeric(N)
sample.size <- nrow(UA_flights)
group.1.size <- nrow(UA_flights[UA_flights$late==1,])
for(i in 1:N)
{
  index <- sample(sample.size, size=group.1.size, replace = FALSE)
  result[i] <- mean(UA_flights$netGain[index], na.rm = TRUE) - mean(UA_flights$netGain[-index], na.rm = TRUE)
}
p <- 2 * (sum(result <= observed) + 1) / (N + 1)
p
```
#### We can see from the p value that it is very less than the standard value of 0.05, so we will reject the null hypothesis and say that there is a difference of average gain between late flights and not late flights
```{r}
observed1 <- avGainVeryLate - avGainNotVeryLate
N <- 10^4 - 1

result1 <- numeric(N)
sample.size <- nrow(UA_flights)
group.1.size <- nrow(UA_flights[UA_flights$very_late==1,])
for(i in 1:N)
{
  index1 <- sample(sample.size, size=group.1.size, replace = FALSE)
  result1[i] <- mean(UA_flights$netGain[index1], na.rm = TRUE) - mean(UA_flights$netGain[-index1], na.rm = TRUE)
}
p <- 2 * (sum(result1 <= observed1) + 1) / (N + 1)
p
```
#### We can see from the p value that it is very less than the standard value of 0.05, so we will reject the null hypothesis and say that there is a difference of average gain between very late flights and not very late flights

```{r}
NYOriginFlights<- c('LGA', 'EWR', 'JFK')

commonFlights <- UA_flights %>% 
  filter(origin %in% NYOriginFlights) %>%
  group_by(dest) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  slice_head(n = 5)

glimpse(commonFlights)
```
```{r}
CommonFlightGain <- subset(UA_flights, dest %in% commonFlights$dest, select = c("dest", "netGain"))
CommonFlightGain %>%
  ggplot(aes(x = netGain,fill=dest)) + 
    geom_bar() +
    facet_wrap(~dest) +
    ggtitle('Net Gain Distribution') +
    xlab('netGain')

```
```{r}
AirDen<-UA_flights %>%
  filter(dest=="DEN")
t.test(AirDen$netGain)$conf
```
```{r}
AirIAH<-UA_flights %>%
  filter(dest=="IAH")
t.test(AirIAH$netGain)$conf
```
```{r}
AirLAX<-UA_flights %>%
  filter(dest=="LAX")
t.test(AirLAX$netGain)$conf
```
```{r}
AirORD<-UA_flights %>%
  filter(dest=="ORD")
t.test(AirORD$netGain)$conf
```
```{r}
AirSFO<-UA_flights %>%
  filter(dest=="SFO")
t.test(AirSFO$netGain)$conf
```


```{r}
CommonFlightGain %>%
  group_by(dest) %>%
  summarize(AvgGain = mean(netGain, na.rm = TRUE))
```
```{r}
UA_flights$DurationHrs<- UA_flights$air_time/60
UA_flights <- UA_flights %>%
  mutate(gainPerHour = netGain/DurationHrs)
```

```{r}
late2 <- UA_flights$late == 1
veryLate2 <- UA_flights$very_late==1
notLate2 <- UA_flights$not_late==1
notveryLate2 <- UA_flights$not_verylate==1
avGainPerHourLate<- mean(UA_flights$gainPerHour[late2])
cat('Average gain per hour for flights that are late: ',avGainPerHourLate,'\n')
avGainPerHourVeryLate<- mean(UA_flights$gainPerHour[veryLate2])
cat('Average gain per hour for flights that are very late: ',avGainPerHourVeryLate,'\n')
avGainPerHourNotLate<- mean(UA_flights$gainPerHour[notLate2])
cat('Average gain per hour for flights that are not late: ',avGainPerHourNotLate,'\n')
avGainPerHourNotVeryLate<- mean(UA_flights$gainPerHour[notveryLate2])
cat('Average gain per hour for flights that are not very late: ',avGainPerHourNotVeryLate,'\n')
```
```{r}
LateFlights1 <-UA_flights %>%
  filter(late==1)
ggplot(data = LateFlights1, mapping = aes(x = gainPerHour)) +
  geom_histogram(color = "white")+
  xlab('Gain Per Hour for Late Flights')
NotLateFlights1 <-UA_flights %>%
  filter(late==0)
ggplot(data = NotLateFlights1, mapping = aes(x = gainPerHour)) +
  geom_histogram(color = "white")+
  xlab('Gain Per Hour for Not Late Flights')
VeryLateFlights1 <-UA_flights %>%
  filter(very_late==1)
ggplot(data = VeryLateFlights1, mapping = aes(x = gainPerHour)) +
  geom_histogram(color = "white")+
  xlab('Gain Per Hour for Very Late Flights')
NotVeryLateFlights1 <-UA_flights %>%
  filter(late==0)
ggplot(data = NotVeryLateFlights1, mapping = aes(x = gainPerHour)) +
  geom_histogram(color = "white")+
  xlab('Gain Per Hour for Not Very Late Flights')
```



```{r}
#### Permutation Tests
observed2 <- avGainPerHourLate - avGainPerHourNotLate
N <- 10^4 - 1

result2 <- numeric(N)
sample.size <- nrow(UA_flights)
group.1.size <- nrow(UA_flights[UA_flights$late==1,])
for(i in 1:N)
{
  index2 <- sample(sample.size, size=group.1.size, replace = FALSE)
  result2[i] <- mean(UA_flights$gainPerHour[index2], na.rm = TRUE) - mean(UA_flights$gainPerHour[-index2], na.rm = TRUE)
}
p <- 2 * (sum(result2 <= observed2) + 1) / (N + 1)
p
```
```{r}
observed3 <- avGainPerHourVeryLate - avGainPerHourNotVeryLate
N <- 10^4 - 1

result3 <- numeric(N)
sample.size <- nrow(UA_flights)
group.1.size <- nrow(UA_flights[UA_flights$very_late==1,])
for(i in 1:N)
{
  index3 <- sample(sample.size, size=group.1.size, replace = FALSE)
  result3[i] <- mean(UA_flights$gainPerHour[index3], na.rm = TRUE) - mean(UA_flights$gainPerHour[-index3], na.rm = TRUE)
}
p <- 2 * (sum(result3 <= observed3) + 1) / (N + 1)
p
```

```{r}
MeanFlightDuration<- mean(UA_flights$air_time)
MeanFlightDuration
UA_flights<- UA_flights %>%
  mutate(l = ifelse(air_time < MeanFlightDuration, 1, 0),
           longFlights = ifelse(air_time >= MeanFlightDuration, 1, 0),
         shortFlights = ifelse(air_time <= MeanFlightDuration, 1, 0))
```
```{r}
ggplot(UA_flights, aes(x = gainPerHour)) +
  geom_histogram(binwidth = 2) +
  facet_wrap(~ shortFlights) +
  labs(title = "Gain per hour Distribution for short flights")
```
```{r}
shortGain <- UA_flights$shortFlights == 1
longGain <- UA_flights$longFlights==1
avGainPerHourShort<- mean(UA_flights$gainPerHour[shortGain])
cat('Average gain per hour for short flights: ',avGainPerHourShort,'\n')
avGainPerHourLong<- mean(UA_flights$gainPerHour[longGain])
cat('Average gain per hour for long flights: ',avGainPerHourLong,'\n')
```

```{r}
#### Permutation Tests
observed4 <- avGainPerHourLong - avGainPerHourShort

N <- 10^4 - 1

result4 <- numeric(N)

sample.size <- nrow(UA_flights)

group.1.size <- nrow(UA_flights[UA_flights$shortFlights==1,])

for(i in 1:N)
{
  index4 <- sample(sample.size, size=group.1.size, replace = FALSE)
  result4[i] <- mean(UA_flights$gainPerHour[index4], na.rm = TRUE) - mean(UA_flights$gainPerHour[-index4], na.rm = TRUE)
}

p <- 2 * (sum(result4 <= observed4) + 1) / (N + 1)
p
```

